The Power of Two Random Choices

### 2014-01-14 article/article.tex
revise sections

The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}. Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details). For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem. Section \ref{sec:multiple-choice} provides an in-depth presentation and analysis of the a basic \emph{multiple-choice balls-into-bins} scheme, the \emph{Greedy} algorithm.

@@

\section{Strategies}
\label{sec:strategies}
We will now consider the most basic variant of the balls-into-bins problem: Namely the static, finite and sequential variant. Thus, the balls are assumed to arrive sequentially, no strategy for deleting balls in the bins has to be considered and we are interested in a bound for the maximum load or gap only after all $m$ balls have arrived.

\subsection{Single-choice balls-into-bins}
\label{sec:single-choice}
In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.

@@

\subsection{Multiple-choice balls-into-bins}
\label{sec:multiple-choice}
The \emph{single-choice balls-into-bins strategy} (Section \ref{sec:single-choice}) chooses one bin for each ball. A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin. We will discuss two different multiple-choice schemes: The \emph{Greedy$[d]$} scheme (Section \ref{sec:greedy}) places the ball into the least full bin among $d$ bins chosen uniformly and independently at random whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a nonuniform selection of $d$ bins.

\subsubsection{The Greedy$[d]$ scheme}
\label{sec:greedy}

@@

The result stated above implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$. This is remarkable in several ways:
\begin{itemize}
\item The apparently small change made for the \emph{multiple-choice balls-into-bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$. 
\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls-into-bins} strategy does not depend on the number of balls $m$. In contrast, the bound for the \emph{one-choice} diverges with $m$.
\item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
\end{itemize}

\subsubsection{The Always-Go-Left scheme}
\label{sec:AlwaysGoLeft}
The multiple-choice balls-into-bins strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random. V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins. Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and independent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}). Furthermore, V\"ocking proved that the \emph{Always-Go-Left} algorithm is almost optimal (TODO: be more precise).

@@

>>>>

The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}. Sections \ref{sec:single-choice} to \ref{sec:multiple-choice} will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details). For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem. Section \ref{sec:multiple-choice} provides an in-depth presentation and analysis of the a basic \emph{multiple-choice balls-into-bins} scheme, the \emph{Greedy} algorithm.

@@

\section{Single-choice balls-into-bins}
\label{sec:single-choice}
We will now consider the most basic variant of the balls-into-bins problem: Namely the static, finite and sequential variant. Thus, the balls are assumed to arrive sequentially, no strategy for deleting balls in the bins has to be considered and we are interested in a bound for the maximum load or gap only after all $m$ balls have arrived.

In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.

@@

\section{Multiple-choice balls-into-bins}
\label{sec:multiple-choice}
The \emph{single-choice balls-into-bins strategy} (Section \ref{sec:single-choice}) chooses one bin for each ball. A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin. We will discuss two different multiple-choice schemes: The \emph{Greedy$[d]$} scheme (Section \ref{sec:greedy}) places the ball into the least full bin among $d$ bins chosen uniformly and independently at random whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a nonuniform selection of $d$ bins.

\subsection{The Greedy$[d]$ scheme}
\label{sec:greedy}

@@

The result stated above implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$. This is remarkable in several ways:
\begin{itemize}
\item The apparently small change made for the \emph{multiple-choice balls-into-bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$. 
\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls-into-bins} strategy does not depend on the number of balls $m$. In contrast, the bound for the \emph{one-choice} diverges with $m$.
\begin{comment}
TODO: this is not true, I think
\item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
\end{comment}
\end{itemize}

\subsection{The Always-Go-Left scheme}
\label{sec:AlwaysGoLeft}
The multiple-choice balls-into-bins strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random. V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins. Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and independent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}). Furthermore, V\"ocking proved that the \emph{Always-Go-Left} algorithm is almost optimal (TODO: be more precise).

@@

\section{Conclusion}
\label{sec:conclusion}
